---
title: "Recommendation Engine"
output: pdf_document
---

## Recommendation Engine - User Collaborative Filtering ##

Recommender systems are active information filtering systems which personalize the information coming to a user based on his interests. These systems are used widely for recommending movies, articles, restaurants, places to visit, items to buy etc.Industry at large uses two different types of recommendation Engines - Content and Collaborative filtering based. This project is all about Collaborative filtering. Further Collaborative filtering is further divided into 2 type - User based and Item based.

This project deals with user based collaborative filtering algorithm. This algorithm usually works by searching a large group of people and finding a smaller set with tastes similar to yours. It looks at other things they like and combines them to create a ranked list of suggestions.

## Library ##
```{r echo = FALSE}
library(reshape2)
library(dtplyr)
library(lsa)
library(SnowballC)
```

## Data ##
```{r echo = FALSE}
setwd("C:/Program Files/RStudio/recomendation_system")
DataMovieRating <- read.csv("movie_rating.csv")
DataMovieRating <- data.frame(DataMovieRating)
```

## Implementing User based Collaborative Filtering ##

This involves two steps: 
Calculating Similarity Function 
Recommend items to users based on user Similarity Score

```{r echo = FALSE}
#data processing and formatting
movie_ratings <- as.data.frame(acast(DataMovieRating, title~critic, value.var="rating"))
```

## Calculate Similarity Scores for Users ##

We need to first fill the N/A where the user has not rated and determine the similarity score. 

```{r echo = FALSE}
x <- movie_ratings[,1:6]
x[is.na(x)] = 0
user_sim <- cosine(as.matrix(t(x)))
```

Creating Similarity score for people helps us to identify similar people


## Recommending ##

For recommending movies we are using the above similarity matrix. As first step, separate the non-rated movies and a weighted matrix is created by multiplying user similarity score (user_sim[,6]) with ratings given by other users.

```{r echo = FALSE}
#create weightge matrix
weight_mat <- user_sim[,6]*movie_ratings[,1:6]
```

Next step is to sum up all the columns of the weight matrix, then divide by the sum of all the similarities for critics that reviewed that movie. The result calculation gives what the user might rate the movie

```{r echo = FALSE}
# Function to recommend items
rec_itm_for_user <- function(userNo)
{
#calcualte column wise sum
col_sums <- list()
rat_user <- movie_ratings[userNo,1:ncol(movie_ratings)]
x <- 1
tot <- list()
z <- 1

for(i in 1:ncol(rat_user)){
  if(is.na(rat_user[1,i]))
  {
    
    col_sums[x] <- sum(weight_mat[,i],na.rm=TRUE)
    x <- x+1
    
    temp <- as.data.frame(weight_mat[,i])
    
    sum_temp = 0
    
    for(j in 1:nrow(temp)){
      if(!is.na(temp[j,1])){
        sum_temp <- sum_temp+user_sim[j,ncol(rat_user)]
      }
      
    }
    tot[z] <- sum_temp
    z <- z+1
  } 
  
  
}
z <- NULL

z <- 1
for(i in 1:ncol(rat_user)){
  if(is.na(rat_user[1,i]))
  {
    rat_user[1,i] <- col_sums[[z]]/tot[[z]]
    z <- z+1
  }
  
}

return(rat_user)
}

#to get N recommendations:
rec_itm_for_user() 

```

## Conclusion ##

So by using the function rec_itm_for_user we can now recommend various movies to each person based on user rating






